% =============================================================
% Block 5 -- Reliability and control (5 slides)
% =============================================================

\sectionframe{Part 5 -- Reliability and control}

% -------- Slide 1: Hallucinations are not only fake citations --------
\begin{frame}{Hallucinations are not only fake citations}
  \vspace{0.2cm}
  \begin{center}
  \begin{tabularx}{0.92\linewidth}{>{\bfseries\color{deepblue}}lX}
  \toprule
  Citation hallucination & A plausible but nonexistent paper, dataset, or table.\\
  Code hallucination & A function that runs, but implements the wrong logic.\\
  Statistical hallucination & Correct syntax, wrong interpretation (e.g., wrong fixed effects).\\
  Workflow hallucination & Task marked complete without real verification (no test, no log, no diff).\\
  \bottomrule
  \end{tabularx}
  \end{center}
  \vspace{0.4cm}
  \midpoint{In agents, the most dangerous hallucination is ``done''.}
  \smallsource{Hallucination taxonomy adapted from Ji et al.\ \citeyearpar{ji_acm_2023}.}
\end{frame}

% -------- Slide 2: Automation bias and overtrust --------
\begin{frame}{Automation bias and overtrust}
  \begin{columns}[T,onlytextwidth]
    \column{0.48\textwidth}
      \threebullets%
        {Humans systematically over-trust automated outputs \citep{skitka_1999_automation_bias}.}%
        {The smoother and more confident the output, the stronger the bias.}%
        {AI outputs are smoother and more confident than human drafts.}
    \column{0.48\textwidth}
      \textbf{What this looks like in practice}
      \begin{itemize}\setlength\itemsep{0.3em}
        \item Accepting a regression specification ``because the output table looks reasonable''.
        \item Copy-pasting a literature summary without reading the cited papers.
        \item Skipping unit tests because the agent ``already checked''.
      \end{itemize}
  \end{columns}
  \vspace{0.4cm}
  \midpoint{Polished prose is not validated reasoning.}
\end{frame}

% -------- Slide 3: Verification by design --------
\begin{frame}{Verification by design}
  \begin{center}
  \begin{tikzpicture}[every node/.style={align=center,font=\small}]
    \node[draw=deepblue,fill=softblue,rounded corners,very thick,
          minimum width=3.2cm,minimum height=1.1cm,text width=3cm]
          (test){\textbf{Tests}\\\scriptsize unit, smoke, golden};
    \node[draw=accent,fill=softorange,rounded corners,very thick,
          minimum width=3.2cm,minimum height=1.1cm,text width=3cm,right=0.7cm of test]
          (diff){\textbf{Diffs}\\\scriptsize visible, reviewable};
    \node[draw=deepblue,fill=softgreen,rounded corners,very thick,
          minimum width=3.2cm,minimum height=1.1cm,text width=3cm,right=0.7cm of diff]
          (logs){\textbf{Logs}\\\scriptsize commands, outputs};
    \node[draw=accent,fill=softpurple,rounded corners,very thick,
          minimum width=3.2cm,minimum height=1.1cm,text width=3cm,right=0.7cm of logs]
          (rev){\textbf{Review}\\\scriptsize human or second agent};
    \draw[->,thick,color=muted] (test)--(diff);
    \draw[->,thick,color=muted] (diff)--(logs);
    \draw[->,thick,color=muted] (logs)--(rev);
  \end{tikzpicture}
  \end{center}
  \vspace{0.4cm}
  \begin{center}\large\color{deepblue}\bfseries Verification cannot be an afterthought; it must be a slot in the workflow.\end{center}
  \vspace{0.2cm}
  \begin{center}\small\color{muted}For research code: at least one test per table, every change as a commit, every command in a log.\end{center}
\end{frame}

% -------- Slide 4: When NOT to use an agent --------
\begin{frame}{When \emph{not} to use an agent}
  \vspace{0.2cm}
  \begin{center}
  \begin{tabularx}{0.92\linewidth}{>{\bfseries\color{deepblue}}lX}
  \toprule
  Identification & The hardest, most causal-inference-specific step.\\
  Confidential data & Where the privacy boundary is unclear.\\
  Final claims & The exact wording of theorems, abstracts, conclusions.\\
  Trust-critical artefacts & Replication packages submitted to a data editor.\\
  Disagreements with coauthors & Resolve by talking, not by prompting.\\
  \bottomrule
  \end{tabularx}
  \end{center}
  \vspace{0.4cm}
  \begin{center}\Large\color{accent}\bfseries Delegating these does not save time; it transfers risk.\end{center}
\end{frame}

% -------- Slide 5: The local alignment problem --------
\begin{frame}{The local alignment problem}
  \begin{center}\Large\bfseries\color{deepblue}The agent optimises for completing the task.\\You care about doing good research.\end{center}
  \vspace{0.4cm}
  \threebullets%
    {Plausible code is not valid code.}%
    {A complete table is not a correct table.}%
    {A polished paragraph is not a justified claim.}
  \vspace{0.5cm}
  \midpoint{The alignment gap is not in 2030; it is in your next prompt.}
\end{frame}
