% =============================================================
% Block 8 -- Closing and bridge (3 slides)
% Yudkowsky/Soares provocation -- 5 take-home messages -- Q&A + sources
% =============================================================

\sectionframe[deepblue]{Part 8 -- Closing}

% -------- Slide 1: Provocation Yudkowsky / Soares --------
\begin{frame}{A provocation worth carrying out of the room}
  \vspace{0.2cm}
  \begin{center}\Large\bfseries\emph{If Anyone Builds It, Everyone Dies.}\end{center}
  \vspace{0.3cm}
  \twobullets%
    {Yudkowsky and Soares \citeyearpar{yudkowsky_soares_2025} offer one of the strongest pessimistic formulations of advanced-AI risk.}%
    {Whether or not you accept the conclusion, the right question to keep is: \emph{what happens when autonomy exceeds control?}}
  \vspace{0.4cm}
  \begin{center}\large\color{deepblue}\bfseries For our daily research, the lesson is modest but firm:\\never delegate without a control structure.\end{center}
\end{frame}

% -------- Slide 2: Five take-home messages --------
\begin{frame}{Five take-home messages}
  \vspace{0.2cm}
  \begin{center}
  \begin{tabularx}{0.94\linewidth}{>{\bfseries\color{accent}\Large}lX}
  \toprule
  1 & Agents are workflows with partial autonomy, not magic assistants.\\
  2 & The right interface depends on the task, not the brand.\\
  3 & Privacy, cost and reproducibility are design choices, not afterthoughts.\\
  4 & Memory is layered: a curated Markdown KB plus a few targeted memory tools beats blind RAG.\\
  5 & The best human role is supervisor, theorist, and final judge.\\
  \bottomrule
  \end{tabularx}
  \end{center}
\end{frame}

% -------- Slide 3: Q&A, contact, sources, update checklist --------
\begin{frame}{Q\&A, contact, sources}
  \begin{columns}[T,onlytextwidth]
    \column{0.50\textwidth}
      \textbf{Contact}\\[0.15cm]
      \small Marco Le Moglie\\
      \small Catholic University of the Sacred Heart\\
      \small \href{https://github.com/MarcoLeMoglie/Agentic_Economic_Research}{\texttt{github.com/MarcoLeMoglie/Agentic\_Economic\_Research}}
      \vspace{0.4cm}
      \textbf{Materials}
      \begin{itemize}\setlength\itemsep{0.2em}
        \item Full slide deck (PDF and LaTeX).
        \item \texttt{bibliography.bib} with every source cited today.
        \item \texttt{README.md} with build instructions.
      \end{itemize}
    \column{0.46\textwidth}
      \textbf{Update checklist (before re-teaching)}
      \begin{itemize}\setlength\itemsep{0.2em}
        \item Prices, plan names, training/privacy defaults.
        \item Adoption numbers (Anthropic Index, Pew, Nature).
        \item Availability of agent features per provider.
        \item Institutional rules (IRB, data editor).
        \item Live demos (URLs and capabilities change).
      \end{itemize}
  \end{columns}
  \vspace{0.5cm}
  \begin{center}\Large\color{deepblue}\bfseries Thank you.\end{center}
\end{frame}

% -------- References slide (auto-generated) --------
\begin{frame}[allowframebreaks]{References}
  \tiny
  \bibliography{bibliography}
\end{frame}
